### 2013-03-15 qrm_tree_gruyere.tex
Ooops, forgot another thing! Shame on me!

% Start of code
% \begin{tikzpicture}[anchor=mid,>=latex',line join=bevel,]
\begin{tikzpicture}[>=latex',line join=bevel,scale=.35]
  \pgfsetlinewidth{1bp}
%%
\pgfsetcolor{black}
  % Edge: 15 -> 31
  \draw [->] (27bp,74.327bp) .. controls (27bp,66.264bp) and (27bp,56.648bp)  .. (27bp,37.667bp);
  % Edge: 9 -> 19
  \draw [->] (459bp,148.33bp) .. controls (459bp,140.26bp) and (459bp,130.65bp)  .. (459bp,111.67bp);
  % Edge: 14 -> 30
  \draw [->] (1050.3bp,151.31bp) .. controls (1060.4bp,140.94bp) and (1073.7bp,127.18bp)  .. (1092bp,108.42bp);
  % Edge: 14 -> 29
  \draw [->] (1035bp,148.33bp) .. controls (1035bp,140.26bp) and (1035bp,130.65bp)  .. (1035bp,111.67bp);
  % Edge: 11 -> 23
  \draw [->] (659.73bp,151.31bp) .. controls (649.64bp,140.94bp) and (636.25bp,127.18bp)  .. (618bp,108.42bp);
  % Edge: 7 -> 15
  \draw [->] (83.731bp,151.31bp) .. controls (73.64bp,140.94bp) and (60.253bp,127.18bp)  .. (42bp,108.42bp);
  % Edge: 0 -> 2
  \draw [->] (588.87bp,377.76bp) .. controls (613.22bp,365.25bp) and (652.76bp,344.93bp)  .. (689.4bp,326.1bp);
  % Edge: 10 -> 22
  \draw [->] (387bp,148.33bp) .. controls (387bp,140.26bp) and (387bp,130.65bp)  .. (387bp,111.67bp);
  % Edge: 13 -> 28
  \draw [->] (935.71bp,149.09bp) .. controls (939.96bp,140.36bp) and (945.15bp,129.68bp)  .. (954.37bp,110.75bp);
  % Edge: 1 -> 3
  \draw [->] (398.64bp,306.66bp) .. controls (360.05bp,293.43bp) and (284.76bp,267.64bp)  .. (231.27bp,249.31bp);
  % Edge: 6 -> 14
  \draw [->] (946.31bp,227.77bp) .. controls (963.36bp,216.09bp) and (988.4bp,198.93bp)  .. (1015.9bp,180.12bp);
  % Edge: 3 -> 8
  \draw [->] (207bp,222.33bp) .. controls (207bp,214.26bp) and (207bp,204.65bp)  .. (207bp,185.67bp);
  % Edge: 13 -> 27
  \draw [->] (918.29bp,149.09bp) .. controls (914.04bp,140.36bp) and (908.85bp,129.68bp)  .. (899.63bp,110.75bp);
  % Edge: 10 -> 21
  \draw [->] (371.73bp,151.31bp) .. controls (361.64bp,140.94bp) and (348.25bp,127.18bp)  .. (330bp,108.42bp);
  % Edge: 12 -> 26
  \draw [->] (762.27bp,151.31bp) .. controls (772.36bp,140.94bp) and (785.75bp,127.18bp)  .. (804bp,108.42bp);
  % Edge: 3 -> 7
  \draw [->] (187.69bp,227.77bp) .. controls (170.64bp,216.09bp) and (145.6bp,198.93bp)  .. (118.14bp,180.12bp);
  % Edge: 5 -> 12
  \draw [->] (719.71bp,223.09bp) .. controls (723.96bp,214.36bp) and (729.15bp,203.68bp)  .. (738.37bp,184.75bp);
  % Edge: 9 -> 20
  \draw [->] (474.27bp,151.31bp) .. controls (484.36bp,140.94bp) and (497.75bp,127.18bp)  .. (516bp,108.42bp);
  % Edge: 1 -> 4
  \draw [->] (423bp,296.33bp) .. controls (423bp,288.26bp) and (423bp,278.65bp)  .. (423bp,259.67bp);
  % Edge: 0 -> 1
  \draw [->] (545.13bp,377.76bp) .. controls (520.78bp,365.25bp) and (481.24bp,344.93bp)  .. (444.6bp,326.1bp);
  % Edge: 2 -> 5
  \draw [->] (711bp,296.33bp) .. controls (711bp,288.26bp) and (711bp,278.65bp)  .. (711bp,259.67bp);
  % Edge: 11 -> 24
  \draw [->] (675bp,148.33bp) .. controls (675bp,140.26bp) and (675bp,130.65bp)  .. (675bp,111.67bp);
  % Edge: 4 -> 10
  \draw [->] (414.29bp,223.09bp) .. controls (410.04bp,214.36bp) and (404.85bp,203.68bp)  .. (395.63bp,184.75bp);
  % Edge: 12 -> 25
  \draw [->] (747bp,148.33bp) .. controls (747bp,140.26bp) and (747bp,130.65bp)  .. (747bp,111.67bp);
  % Edge: 8 -> 17
  \draw [->] (198.29bp,149.09bp) .. controls (194.04bp,140.36bp) and (188.85bp,129.68bp)  .. (179.63bp,110.75bp);
  % Edge: 6 -> 13
  \draw [->] (927bp,222.33bp) .. controls (927bp,214.26bp) and (927bp,204.65bp)  .. (927bp,185.67bp);
  % Edge: 5 -> 11
  \draw [->] (702.29bp,223.09bp) .. controls (698.04bp,214.36bp) and (692.85bp,203.68bp)  .. (683.63bp,184.75bp);
  % Edge: 7 -> 16
  \draw [->] (99bp,148.33bp) .. controls (99bp,140.26bp) and (99bp,130.65bp)  .. (99bp,111.67bp);
  % Edge: 2 -> 6
  \draw [->] (735.36bp,306.66bp) .. controls (773.95bp,293.43bp) and (849.24bp,267.64bp)  .. (902.73bp,249.31bp);
  % Edge: 8 -> 18
  \draw [->] (215.71bp,149.09bp) .. controls (219.96bp,140.36bp) and (225.15bp,129.68bp)  .. (234.37bp,110.75bp);
  % Edge: 4 -> 9
  \draw [->] (431.71bp,223.09bp) .. controls (435.96bp,214.36bp) and (441.15bp,203.68bp)  .. (450.37bp,184.75bp);
  % Node: 24
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (675bp,93bp) ellipse (27bp and 18bp);
  \draw (675bp,93bp) node {24};
\end{scope}
  % Node: 25
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (747bp,93bp) ellipse (27bp and 18bp);
  \draw (747bp,93bp) node {25};
\end{scope}
  % Node: 26
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (819bp,93bp) ellipse (27bp and 18bp);
  \draw (819bp,93bp) node {26};
\end{scope}
  % Node: 27
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (891bp,93bp) ellipse (27bp and 18bp);
  \draw (891bp,93bp) node {27};
\end{scope}
  % Node: 20
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (531bp,93bp) ellipse (27bp and 18bp);
  \draw (531bp,93bp) node {20};
\end{scope}
  % Node: 21
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (315bp,93bp) ellipse (27bp and 18bp);
  \draw (315bp,93bp) node {21};
\end{scope}
  % Node: 22
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (387bp,93bp) ellipse (27bp and 18bp);
  \draw (387bp,93bp) node {22};
\end{scope}
  % Node: 23
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (603bp,93bp) ellipse (27bp and 18bp);
  \draw (603bp,93bp) node {23};
\end{scope}
  % Node: 28
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (963bp,93bp) ellipse (27bp and 18bp);
  \draw (963bp,93bp) node {28};
\end{scope}
  % Node: 29
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (1035bp,93bp) ellipse (27bp and 18bp);
  \draw (1035bp,93bp) node {29};
\end{scope}
  % Node: 1
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (423bp,315bp) ellipse (27bp and 18bp);
  \draw (423bp,315bp) node {1};
\end{scope}
  % Node: 0
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (567bp,389bp) ellipse (27bp and 18bp);
  \draw (567bp,389bp) node {0};
\end{scope}
  % Node: 3
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (207bp,241bp) ellipse (27bp and 18bp);
  \draw (207bp,241bp) node {3};
\end{scope}
  % Node: 2
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (711bp,315bp) ellipse (27bp and 18bp);
  \draw (711bp,315bp) node {2};
\end{scope}
  % Node: 5
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (711bp,241bp) ellipse (27bp and 18bp);
  \draw (711bp,241bp) node {5};
\end{scope}
  % Node: 4
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (423bp,241bp) ellipse (27bp and 18bp);
  \draw (423bp,241bp) node {4};
\end{scope}
  % Node: 7
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (99bp,167bp) ellipse (27bp and 18bp);
  \draw (99bp,167bp) node {7};
\end{scope}
  % Node: 6
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (927bp,241bp) ellipse (27bp and 18bp);
  \draw (927bp,241bp) node {6};
\end{scope}
  % Node: 9
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (459bp,167bp) ellipse (27bp and 18bp);
  \draw (459bp,167bp) node {9};
\end{scope}
  % Node: 8
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (207bp,167bp) ellipse (27bp and 18bp);
  \draw (207bp,167bp) node {8};
\end{scope}
  % Node: 11
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (675bp,167bp) ellipse (27bp and 18bp);
  \draw (675bp,167bp) node {11};
\end{scope}
  % Node: 10
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (387bp,167bp) ellipse (27bp and 18bp);
  \draw (387bp,167bp) node {10};
\end{scope}
  % Node: 13
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (927bp,167bp) ellipse (27bp and 18bp);
  \draw (927bp,167bp) node {13};
\end{scope}
  % Node: 12
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (747bp,167bp) ellipse (27bp and 18bp);
  \draw (747bp,167bp) node {12};
\end{scope}
  % Node: 15
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (27bp,93bp) ellipse (27bp and 18bp);
  \draw (27bp,93bp) node {15};
\end{scope}
  % Node: 14
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (1035bp,167bp) ellipse (27bp and 18bp);
  \draw (1035bp,167bp) node {14};
\end{scope}
  % Node: 17
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (171bp,93bp) ellipse (27bp and 18bp);
  \draw (171bp,93bp) node {17};
\end{scope}
  % Node: 16
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (99bp,93bp) ellipse (27bp and 18bp);
  \draw (99bp,93bp) node {16};
\end{scope}
  % Node: 19
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (459bp,93bp) ellipse (27bp and 18bp);
  \draw (459bp,93bp) node {19};
\end{scope}
  % Node: 18
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (243bp,93bp) ellipse (27bp and 18bp);
  \draw (243bp,93bp) node {18};
\end{scope}
  % Node: 31
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (27bp,19bp) ellipse (27bp and 18bp);
  \draw (27bp,19bp) node {31};
\end{scope}
  % Node: 30
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (1107bp,93bp) ellipse (27bp and 18bp);
  \draw (1107bp,93bp) node {30};
\end{scope}
%
\end{tikzpicture}
% End of code