### 2013-05-07 visu_hierarchy.tex
Oops some files missing

\node at (0mm,0mm) {core 00};
\draw[color=black!10] (20mm,0mm) -- (30cm,0mm);
\node at (0mm,15mm) {core 01};
\draw[color=black!10] (20mm,15mm) -- (30cm,15mm);
\node at (0mm,30mm) {core 02};
\draw[color=black!10] (20mm,30mm) -- (30cm,30mm);
\node at (0mm,45mm) {core 03};
\draw[color=black!10] (20mm,45mm) -- (30cm,45mm);
\node at (0mm,60mm) {core 04};
\draw[color=black!10] (20mm,60mm) -- (30cm,60mm);
\node at (0mm,75mm) {core 05};
\draw[color=black!10] (20mm,75mm) -- (30cm,75mm);
\node at (0mm,90mm) {core 06};
\draw[color=black!10] (20mm,90mm) -- (30cm,90mm);
\node at (0mm,105mm) {core 07};
\draw[color=black!10] (20mm,105mm) -- (30cm,105mm);
\node at (0mm,120mm) {core 08};
\draw[color=black!10] (20mm,120mm) -- (30cm,120mm);
\node at (0mm,135mm) {core 09};
\draw[color=black!10] (20mm,135mm) -- (30cm,135mm);
\node at (0mm,150mm) {core 10};
\draw[color=black!10] (20mm,150mm) -- (30cm,150mm);
\node at (0mm,165mm) {core 11};
\draw[color=black!10] (20mm,165mm) -- (30cm,165mm);
\node at (0mm,180mm) {core 12};
\draw[color=black!10] (20mm,180mm) -- (30cm,180mm);
\node at (0mm,195mm) {core 13};
\draw[color=black!10] (20mm,195mm) -- (30cm,195mm);
\node at (0mm,210mm) {core 14};
\draw[color=black!10] (20mm,210mm) -- (30cm,210mm);
\node at (0mm,225mm) {core 15};
\draw[color=black!10] (20mm,225mm) -- (30cm,225mm);
\node at (0mm,240mm) {core 16};
\draw[color=black!10] (20mm,240mm) -- (30cm,240mm);
\node at (0mm,255mm) {core 17};
\draw[color=black!10] (20mm,255mm) -- (30cm,255mm);
\node at (0mm,270mm) {core 18};
\draw[color=black!10] (20mm,270mm) -- (30cm,270mm);
\node at (0mm,285mm) {core 19};
\draw[color=black!10] (20mm,285mm) -- (30cm,285mm);
\node at (0mm,300mm) {core 20};
\draw[color=black!10] (20mm,300mm) -- (30cm,300mm);
\node at (0mm,315mm) {core 21};
\draw[color=black!10] (20mm,315mm) -- (30cm,315mm);
\node at (0mm,330mm) {core 22};
\draw[color=black!10] (20mm,330mm) -- (30cm,330mm);
\node at (0mm,345mm) {core 23};
\draw[color=black!10] (20mm,345mm) -- (30cm,345mm);
\node at (0mm,360mm) {core 24};
\draw[color=black!10] (20mm,360mm) -- (30cm,360mm);
\node at (0mm,375mm) {core 25};
\draw[color=black!10] (20mm,375mm) -- (30cm,375mm);
\node at (0mm,390mm) {core 26};
\draw[color=black!10] (20mm,390mm) -- (30cm,390mm);
\node at (0mm,405mm) {core 27};
\draw[color=black!10] (20mm,405mm) -- (30cm,405mm);
\node at (0mm,420mm) {core 28};
\draw[color=black!10] (20mm,420mm) -- (30cm,420mm);
\node at (0mm,435mm) {core 29};
\draw[color=black!10] (20mm,435mm) -- (30cm,435mm);
\node at (0mm,450mm) {core 30};
\draw[color=black!10] (20mm,450mm) -- (30cm,450mm);
\node at (0mm,465mm) {core 31};
\draw[color=black!10] (20mm,465mm) -- (30cm,465mm);
\node[draw,fill=red!20,minimum size=10mm] (s_0_8) at (50mm,0mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_0_4) at (60mm,0mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_0_3) at (70mm,0mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_0_2) at (80mm,0mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_0_1) at (90mm,0mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_0_8) at (80mm,120mm) {};
\draw[->] (s_0_8.center) -- (r_0_8.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_8_20) at (90mm,120mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_8_24) at (100mm,120mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_8_10) at (110mm,120mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_8_9) at (120mm,120mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_8_11) at (130mm,120mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_0_4) at (80mm,60mm) {};
\draw[->] (s_0_4.center) -- (r_0_4.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_4_12) at (90mm,60mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_4_16) at (100mm,60mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_4_7) at (110mm,60mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_4_6) at (120mm,60mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_4_5) at (130mm,60mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_0_3) at (81mm,45mm) {};
\draw[->] (s_0_3.center) -- (r_0_3.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_0_2) at (91mm,30mm) {};
\draw[->] (s_0_2.center) -- (r_0_2.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_0_1) at (101mm,15mm) {};
\draw[->] (s_0_1.center) -- (r_0_1.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_8_20) at (120mm,300mm) {};
\draw[->] (s_8_20.center) -- (r_8_20.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_20_21) at (130mm,300mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_20_22) at (140mm,300mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_20_23) at (150mm,300mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_8_24) at (120mm,360mm) {};
\draw[->] (s_8_24.center) -- (r_8_24.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_24_26) at (130mm,360mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_24_27) at (140mm,360mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_24_25) at (150mm,360mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_4_12) at (120mm,180mm) {};
\draw[->] (s_4_12.center) -- (r_4_12.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_12_28) at (130mm,180mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_12_15) at (140mm,180mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_12_14) at (150mm,180mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_12_13) at (160mm,180mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_8_10) at (121mm,150mm) {};
\draw[->] (s_8_10.center) -- (r_8_10.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_4_7) at (121mm,105mm) {};
\draw[->] (s_4_7.center) -- (r_4_7.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_4_16) at (130mm,240mm) {};
\draw[->] (s_4_16.center) -- (r_4_16.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_16_17) at (140mm,240mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_16_18) at (150mm,240mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_16_19) at (160mm,240mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_8_9) at (131mm,135mm) {};
\draw[->] (s_8_9.center) -- (r_8_9.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_4_6) at (131mm,90mm) {};
\draw[->] (s_4_6.center) -- (r_4_6.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_4_5) at (141mm,75mm) {};
\draw[->] (s_4_5.center) -- (r_4_5.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_8_11) at (141mm,165mm) {};
\draw[->] (s_8_11.center) -- (r_8_11.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_24_26) at (141mm,390mm) {};
\draw[->] (s_24_26.center) -- (r_24_26.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_20_21) at (141mm,315mm) {};
\draw[->] (s_20_21.center) -- (r_20_21.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_12_28) at (150mm,420mm) {};
\draw[->] (s_12_28.center) -- (r_12_28.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_28_31) at (160mm,420mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_28_29) at (170mm,420mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_28_30) at (180mm,420mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_16_17) at (151mm,255mm) {};
\draw[->] (s_16_17.center) -- (r_16_17.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_20_22) at (151mm,330mm) {};
\draw[->] (s_20_22.center) -- (r_20_22.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_24_27) at (151mm,405mm) {};
\draw[->] (s_24_27.center) -- (r_24_27.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_12_15) at (151mm,225mm) {};
\draw[->] (s_12_15.center) -- (r_12_15.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_20_23) at (161mm,345mm) {};
\draw[->] (s_20_23.center) -- (r_20_23.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_24_25) at (161mm,375mm) {};
\draw[->] (s_24_25.center) -- (r_24_25.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_12_14) at (161mm,210mm) {};
\draw[->] (s_12_14.center) -- (r_12_14.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_16_18) at (161mm,270mm) {};
\draw[->] (s_16_18.center) -- (r_16_18.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_16_19) at (171mm,285mm) {};
\draw[->] (s_16_19.center) -- (r_16_19.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_28_31) at (171mm,465mm) {};
\draw[->] (s_28_31.center) -- (r_28_31.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_12_13) at (171mm,195mm) {};
\draw[->] (s_12_13.center) -- (r_12_13.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_28_29) at (181mm,435mm) {};
\draw[->] (s_28_29.center) -- (r_28_29.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_28_30) at (191mm,450mm) {};
\draw[->] (s_28_30.center) -- (r_28_30.center);