### 2013-06-04 graphs/visu_gruyere_badtree.tex
Added forgotten file

\draw[fill,color=red!10] (-3cm,7.500000mm) rectangle (30cm,-7.500000mm);
\node at (0mm,0mm) {core 00};
\draw[color=black!30] (20mm,0mm) -- (30cm,0mm);
\draw[fill,color=red!10] (-3cm,22.500000mm) rectangle (30cm,7.500000mm);
\node at (0mm,15mm) {core 01};
\draw[color=black!30] (20mm,15mm) -- (30cm,15mm);
\draw[fill,color=red!10] (-3cm,37.500000mm) rectangle (30cm,22.500000mm);
\node at (0mm,30mm) {core 02};
\draw[color=black!30] (20mm,30mm) -- (30cm,30mm);
\draw[fill,color=red!10] (-3cm,52.500000mm) rectangle (30cm,37.500000mm);
\node at (0mm,45mm) {core 03};
\draw[color=black!30] (20mm,45mm) -- (30cm,45mm);
\draw[fill,color=green!10] (-3cm,67.500000mm) rectangle (30cm,52.500000mm);
\node at (0mm,60mm) {core 04};
\draw[color=black!30] (20mm,60mm) -- (30cm,60mm);
\draw[fill,color=green!10] (-3cm,82.500000mm) rectangle (30cm,67.500000mm);
\node at (0mm,75mm) {core 05};
\draw[color=black!30] (20mm,75mm) -- (30cm,75mm);
\draw[fill,color=green!10] (-3cm,97.500000mm) rectangle (30cm,82.500000mm);
\node at (0mm,90mm) {core 06};
\draw[color=black!30] (20mm,90mm) -- (30cm,90mm);
\draw[fill,color=green!10] (-3cm,112.500000mm) rectangle (30cm,97.500000mm);
\node at (0mm,105mm) {core 07};
\draw[color=black!30] (20mm,105mm) -- (30cm,105mm);
\draw[fill,color=blue!10] (-3cm,127.500000mm) rectangle (30cm,112.500000mm);
\node at (0mm,120mm) {core 08};
\draw[color=black!30] (20mm,120mm) -- (30cm,120mm);
\draw[fill,color=blue!10] (-3cm,142.500000mm) rectangle (30cm,127.500000mm);
\node at (0mm,135mm) {core 09};
\draw[color=black!30] (20mm,135mm) -- (30cm,135mm);
\draw[fill,color=blue!10] (-3cm,157.500000mm) rectangle (30cm,142.500000mm);
\node at (0mm,150mm) {core 10};
\draw[color=black!30] (20mm,150mm) -- (30cm,150mm);
\draw[fill,color=blue!10] (-3cm,172.500000mm) rectangle (30cm,157.500000mm);
\node at (0mm,165mm) {core 11};
\draw[color=black!30] (20mm,165mm) -- (30cm,165mm);
\draw[fill,color=orange!10] (-3cm,187.500000mm) rectangle (30cm,172.500000mm);
\node at (0mm,180mm) {core 12};
\draw[color=black!30] (20mm,180mm) -- (30cm,180mm);
\draw[fill,color=orange!10] (-3cm,202.500000mm) rectangle (30cm,187.500000mm);
\node at (0mm,195mm) {core 13};
\draw[color=black!30] (20mm,195mm) -- (30cm,195mm);
\draw[fill,color=orange!10] (-3cm,217.500000mm) rectangle (30cm,202.500000mm);
\node at (0mm,210mm) {core 14};
\draw[color=black!30] (20mm,210mm) -- (30cm,210mm);
\draw[fill,color=orange!10] (-3cm,232.500000mm) rectangle (30cm,217.500000mm);
\node at (0mm,225mm) {core 15};
\draw[color=black!30] (20mm,225mm) -- (30cm,225mm);
\draw[fill,color=red!10] (-3cm,247.500000mm) rectangle (30cm,232.500000mm);
\node at (0mm,240mm) {core 16};
\draw[color=black!30] (20mm,240mm) -- (30cm,240mm);
\draw[fill,color=red!10] (-3cm,262.500000mm) rectangle (30cm,247.500000mm);
\node at (0mm,255mm) {core 17};
\draw[color=black!30] (20mm,255mm) -- (30cm,255mm);
\draw[fill,color=red!10] (-3cm,277.500000mm) rectangle (30cm,262.500000mm);
\node at (0mm,270mm) {core 18};
\draw[color=black!30] (20mm,270mm) -- (30cm,270mm);
\draw[fill,color=red!10] (-3cm,292.500000mm) rectangle (30cm,277.500000mm);
\node at (0mm,285mm) {core 19};
\draw[color=black!30] (20mm,285mm) -- (30cm,285mm);
\draw[fill,color=green!10] (-3cm,307.500000mm) rectangle (30cm,292.500000mm);
\node at (0mm,300mm) {core 20};
\draw[color=black!30] (20mm,300mm) -- (30cm,300mm);
\draw[fill,color=green!10] (-3cm,322.500000mm) rectangle (30cm,307.500000mm);
\node at (0mm,315mm) {core 21};
\draw[color=black!30] (20mm,315mm) -- (30cm,315mm);
\draw[fill,color=green!10] (-3cm,337.500000mm) rectangle (30cm,322.500000mm);
\node at (0mm,330mm) {core 22};
\draw[color=black!30] (20mm,330mm) -- (30cm,330mm);
\draw[fill,color=green!10] (-3cm,352.500000mm) rectangle (30cm,337.500000mm);
\node at (0mm,345mm) {core 23};
\draw[color=black!30] (20mm,345mm) -- (30cm,345mm);
\draw[fill,color=blue!10] (-3cm,367.500000mm) rectangle (30cm,352.500000mm);
\node at (0mm,360mm) {core 24};
\draw[color=black!30] (20mm,360mm) -- (30cm,360mm);
\draw[fill,color=blue!10] (-3cm,382.500000mm) rectangle (30cm,367.500000mm);
\node at (0mm,375mm) {core 25};
\draw[color=black!30] (20mm,375mm) -- (30cm,375mm);
\draw[fill,color=blue!10] (-3cm,397.500000mm) rectangle (30cm,382.500000mm);
\node at (0mm,390mm) {core 26};
\draw[color=black!30] (20mm,390mm) -- (30cm,390mm);
\draw[fill,color=blue!10] (-3cm,412.500000mm) rectangle (30cm,397.500000mm);
\node at (0mm,405mm) {core 27};
\draw[color=black!30] (20mm,405mm) -- (30cm,405mm);
\draw[fill,color=orange!10] (-3cm,427.500000mm) rectangle (30cm,412.500000mm);
\node at (0mm,420mm) {core 28};
\draw[color=black!30] (20mm,420mm) -- (30cm,420mm);
\draw[fill,color=orange!10] (-3cm,442.500000mm) rectangle (30cm,427.500000mm);
\node at (0mm,435mm) {core 29};
\draw[color=black!30] (20mm,435mm) -- (30cm,435mm);
\draw[fill,color=orange!10] (-3cm,457.500000mm) rectangle (30cm,442.500000mm);
\node at (0mm,450mm) {core 30};
\draw[color=black!30] (20mm,450mm) -- (30cm,450mm);
\draw[fill,color=orange!10] (-3cm,472.500000mm) rectangle (30cm,457.500000mm);
\node at (0mm,465mm) {core 31};
\draw[color=black!30] (20mm,465mm) -- (30cm,465mm);
\node at (0mm,480mm) {Machine: gruyere, topology: badtree};
\node[draw,fill=red!20,minimum size=10mm] (s_0_14) at (50mm,0mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_0_28) at (60mm,0mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_0_14) at (90mm,210mm) {};
\draw[->,very thick,color=red] (s_0_14.center) -- (r_0_14.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_14_1) at (100mm,210mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_14_19) at (110mm,210mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_0_28) at (100mm,420mm) {};
\draw[->,very thick,color=red] (s_0_28.center) -- (r_0_28.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_28_3) at (110mm,420mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_28_17) at (120mm,420mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_14_1) at (140mm,15mm) {};
\draw[->,very thick,color=red] (s_14_1.center) -- (r_14_1.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_1_27) at (150mm,15mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_1_15) at (160mm,15mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_14_19) at (150mm,285mm) {};
\draw[->,very thick,color=red] (s_14_19.center) -- (r_14_19.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_19_13) at (160mm,285mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_19_9) at (170mm,285mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_28_3) at (150mm,45mm) {};
\draw[->,very thick,color=red] (s_28_3.center) -- (r_28_3.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_3_26) at (160mm,45mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_3_25) at (170mm,45mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_28_17) at (160mm,255mm) {};
\draw[->,very thick,color=red] (s_28_17.center) -- (r_28_17.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_17_29) at (170mm,255mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_17_12) at (180mm,255mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_1_27) at (190mm,405mm) {};
\draw[->,very thick,color=red] (s_1_27.center) -- (r_1_27.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_27_2) at (200mm,405mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_27_6) at (210mm,405mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_3_26) at (200mm,390mm) {};
\draw[->,very thick,color=red] (s_3_26.center) -- (r_3_26.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_26_5) at (210mm,390mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_19_13) at (200mm,195mm) {};
\draw[->,very thick,color=red] (s_19_13.center) -- (r_19_13.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_13_21) at (210mm,195mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_13_20) at (220mm,195mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_1_15) at (200mm,225mm) {};
\draw[->,very thick,color=red] (s_1_15.center) -- (r_1_15.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_15_18) at (210mm,225mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_15_16) at (220mm,225mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_3_25) at (210mm,375mm) {};
\draw[->,very thick,color=red] (s_3_25.center) -- (r_3_25.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_25_7) at (220mm,375mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_25_4) at (230mm,375mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_19_9) at (210mm,135mm) {};
\draw[->,very thick,color=red] (s_19_9.center) -- (r_19_9.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_9_23) at (220mm,135mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_9_22) at (230mm,135mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_17_29) at (210mm,435mm) {};
\draw[->,very thick,color=red] (s_17_29.center) -- (r_17_29.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_17_12) at (220mm,180mm) {};
\draw[->,very thick,color=red] (s_17_12.center) -- (r_17_12.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_27_2) at (240mm,30mm) {};
\draw[->,very thick,color=red] (s_27_2.center) -- (r_27_2.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_2_30) at (250mm,30mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_2_24) at (260mm,30mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_26_5) at (240mm,75mm) {};
\draw[->,very thick,color=red] (s_26_5.center) -- (r_26_5.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_27_6) at (240mm,90mm) {};
\draw[->,very thick,color=red] (s_27_6.center) -- (r_27_6.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_13_21) at (240mm,315mm) {};
\draw[->,very thick,color=red] (s_13_21.center) -- (r_13_21.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_13_20) at (250mm,300mm) {};
\draw[->,very thick,color=red] (s_13_20.center) -- (r_13_20.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_15_18) at (250mm,270mm) {};
\draw[->,very thick,color=red] (s_15_18.center) -- (r_15_18.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_18_10) at (260mm,270mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_18_8) at (270mm,270mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_25_7) at (250mm,105mm) {};
\draw[->,very thick,color=red] (s_25_7.center) -- (r_25_7.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_9_23) at (250mm,345mm) {};
\draw[->,very thick,color=red] (s_9_23.center) -- (r_9_23.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_9_22) at (260mm,330mm) {};
\draw[->,very thick,color=red] (s_9_22.center) -- (r_9_22.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_25_4) at (260mm,60mm) {};
\draw[->,very thick,color=red] (s_25_4.center) -- (r_25_4.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_15_16) at (260mm,240mm) {};
\draw[->,very thick,color=red] (s_15_16.center) -- (r_15_16.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_16_31) at (270mm,240mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_16_11) at (280mm,240mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_2_30) at (290mm,450mm) {};
\draw[->,very thick,color=red] (s_2_30.center) -- (r_2_30.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_2_24) at (300mm,360mm) {};
\draw[->,very thick,color=red] (s_2_24.center) -- (r_2_24.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_18_10) at (300mm,150mm) {};
\draw[->,very thick,color=red] (s_18_10.center) -- (r_18_10.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_18_8) at (310mm,120mm) {};
\draw[->,very thick,color=red] (s_18_8.center) -- (r_18_8.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_16_31) at (310mm,465mm) {};
\draw[->,very thick,color=red] (s_16_31.center) -- (r_16_31.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_16_11) at (320mm,165mm) {};
\draw[->,very thick,color=red] (s_16_11.center) -- (r_16_11.center);