### 2013-06-04 graphs/visu_ziger_mst.tex

\draw[fill,color=red!10] (-3cm,7.500000mm) rectangle (30cm,-7.500000mm);
\node at (0mm,0mm) {core 00};
\draw[color=black!30] (20mm,0mm) -- (30cm,0mm);
\draw[fill,color=red!10] (-3cm,22.500000mm) rectangle (30cm,7.500000mm);
\node at (0mm,15mm) {core 01};
\draw[color=black!30] (20mm,15mm) -- (30cm,15mm);
\draw[fill,color=red!10] (-3cm,37.500000mm) rectangle (30cm,22.500000mm);
\node at (0mm,30mm) {core 02};
\draw[color=black!30] (20mm,30mm) -- (30cm,30mm);
\draw[fill,color=red!10] (-3cm,52.500000mm) rectangle (30cm,37.500000mm);
\node at (0mm,45mm) {core 03};
\draw[color=black!30] (20mm,45mm) -- (30cm,45mm);
\draw[fill,color=red!10] (-3cm,67.500000mm) rectangle (30cm,52.500000mm);
\node at (0mm,60mm) {core 04};
\draw[color=black!30] (20mm,60mm) -- (30cm,60mm);
\draw[fill,color=red!10] (-3cm,82.500000mm) rectangle (30cm,67.500000mm);
\node at (0mm,75mm) {core 05};
\draw[color=black!30] (20mm,75mm) -- (30cm,75mm);
\draw[fill,color=green!10] (-3cm,97.500000mm) rectangle (30cm,82.500000mm);
\node at (0mm,90mm) {core 06};
\draw[color=black!30] (20mm,90mm) -- (30cm,90mm);
\draw[fill,color=green!10] (-3cm,112.500000mm) rectangle (30cm,97.500000mm);
\node at (0mm,105mm) {core 07};
\draw[color=black!30] (20mm,105mm) -- (30cm,105mm);
\draw[fill,color=green!10] (-3cm,127.500000mm) rectangle (30cm,112.500000mm);
\node at (0mm,120mm) {core 08};
\draw[color=black!30] (20mm,120mm) -- (30cm,120mm);
\draw[fill,color=green!10] (-3cm,142.500000mm) rectangle (30cm,127.500000mm);
\node at (0mm,135mm) {core 09};
\draw[color=black!30] (20mm,135mm) -- (30cm,135mm);
\draw[fill,color=green!10] (-3cm,157.500000mm) rectangle (30cm,142.500000mm);
\node at (0mm,150mm) {core 10};
\draw[color=black!30] (20mm,150mm) -- (30cm,150mm);
\draw[fill,color=green!10] (-3cm,172.500000mm) rectangle (30cm,157.500000mm);
\node at (0mm,165mm) {core 11};
\draw[color=black!30] (20mm,165mm) -- (30cm,165mm);
\draw[fill,color=blue!10] (-3cm,187.500000mm) rectangle (30cm,172.500000mm);
\node at (0mm,180mm) {core 12};
\draw[color=black!30] (20mm,180mm) -- (30cm,180mm);
\draw[fill,color=blue!10] (-3cm,202.500000mm) rectangle (30cm,187.500000mm);
\node at (0mm,195mm) {core 13};
\draw[color=black!30] (20mm,195mm) -- (30cm,195mm);
\draw[fill,color=blue!10] (-3cm,217.500000mm) rectangle (30cm,202.500000mm);
\node at (0mm,210mm) {core 14};
\draw[color=black!30] (20mm,210mm) -- (30cm,210mm);
\draw[fill,color=blue!10] (-3cm,232.500000mm) rectangle (30cm,217.500000mm);
\node at (0mm,225mm) {core 15};
\draw[color=black!30] (20mm,225mm) -- (30cm,225mm);
\draw[fill,color=blue!10] (-3cm,247.500000mm) rectangle (30cm,232.500000mm);
\node at (0mm,240mm) {core 16};
\draw[color=black!30] (20mm,240mm) -- (30cm,240mm);
\draw[fill,color=blue!10] (-3cm,262.500000mm) rectangle (30cm,247.500000mm);
\node at (0mm,255mm) {core 17};
\draw[color=black!30] (20mm,255mm) -- (30cm,255mm);
\draw[fill,color=orange!10] (-3cm,277.500000mm) rectangle (30cm,262.500000mm);
\node at (0mm,270mm) {core 18};
\draw[color=black!30] (20mm,270mm) -- (30cm,270mm);
\draw[fill,color=orange!10] (-3cm,292.500000mm) rectangle (30cm,277.500000mm);
\node at (0mm,285mm) {core 19};
\draw[color=black!30] (20mm,285mm) -- (30cm,285mm);
\draw[fill,color=orange!10] (-3cm,307.500000mm) rectangle (30cm,292.500000mm);
\node at (0mm,300mm) {core 20};
\draw[color=black!30] (20mm,300mm) -- (30cm,300mm);
\draw[fill,color=orange!10] (-3cm,322.500000mm) rectangle (30cm,307.500000mm);
\node at (0mm,315mm) {core 21};
\draw[color=black!30] (20mm,315mm) -- (30cm,315mm);
\draw[fill,color=orange!10] (-3cm,337.500000mm) rectangle (30cm,322.500000mm);
\node at (0mm,330mm) {core 22};
\draw[color=black!30] (20mm,330mm) -- (30cm,330mm);
\draw[fill,color=orange!10] (-3cm,352.500000mm) rectangle (30cm,337.500000mm);
\node at (0mm,345mm) {core 23};
\draw[color=black!30] (20mm,345mm) -- (30cm,345mm);
\node at (0mm,360mm) {Machine: ziger, topology: mst};
\node[draw,fill=red!20,minimum size=10mm] (s_0_6) at (50mm,0mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_0_18) at (60mm,0mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_0_5) at (70mm,0mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_0_4) at (80mm,0mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_0_3) at (90mm,0mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_0_2) at (100mm,0mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_0_1) at (110mm,0mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_0_6) at (70mm,90mm) {};
\draw[->,very thick,color=red] (s_0_6.center) -- (r_0_6.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_6_12) at (80mm,90mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_6_11) at (90mm,90mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_6_10) at (100mm,90mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_6_9) at (110mm,90mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_6_8) at (120mm,90mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_6_7) at (130mm,90mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_0_18) at (80mm,270mm) {};
\draw[->,very thick,color=red] (s_0_18.center) -- (r_0_18.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_18_23) at (90mm,270mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_18_22) at (100mm,270mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_18_21) at (110mm,270mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_18_20) at (120mm,270mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_18_19) at (130mm,270mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_0_5) at (81mm,75mm) {};
\draw[->] (s_0_5.center) -- (r_0_5.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_0_4) at (91mm,60mm) {};
\draw[->] (s_0_4.center) -- (r_0_4.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_6_12) at (100mm,180mm) {};
\draw[->,very thick,color=red] (s_6_12.center) -- (r_6_12.center); 
\node[draw,fill=red!20,minimum size=10mm] (s_12_17) at (110mm,180mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_12_16) at (120mm,180mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_12_15) at (130mm,180mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_12_14) at (140mm,180mm) {};
\node[draw,fill=red!20,minimum size=10mm] (s_12_13) at (150mm,180mm) {};
\node[draw,fill=blue!20,minimum size=10mm] (r_6_11) at (101mm,165mm) {};
\draw[->] (s_6_11.center) -- (r_6_11.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_0_3) at (101mm,45mm) {};
\draw[->] (s_0_3.center) -- (r_0_3.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_18_23) at (101mm,345mm) {};
\draw[->] (s_18_23.center) -- (r_18_23.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_0_2) at (111mm,30mm) {};
\draw[->] (s_0_2.center) -- (r_0_2.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_18_22) at (111mm,330mm) {};
\draw[->] (s_18_22.center) -- (r_18_22.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_6_10) at (111mm,150mm) {};
\draw[->] (s_6_10.center) -- (r_6_10.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_6_9) at (121mm,135mm) {};
\draw[->] (s_6_9.center) -- (r_6_9.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_0_1) at (121mm,15mm) {};
\draw[->] (s_0_1.center) -- (r_0_1.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_12_17) at (121mm,255mm) {};
\draw[->] (s_12_17.center) -- (r_12_17.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_18_21) at (121mm,315mm) {};
\draw[->] (s_18_21.center) -- (r_18_21.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_6_8) at (131mm,120mm) {};
\draw[->] (s_6_8.center) -- (r_6_8.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_12_16) at (131mm,240mm) {};
\draw[->] (s_12_16.center) -- (r_12_16.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_18_20) at (131mm,300mm) {};
\draw[->] (s_18_20.center) -- (r_18_20.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_6_7) at (141mm,105mm) {};
\draw[->] (s_6_7.center) -- (r_6_7.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_18_19) at (141mm,285mm) {};
\draw[->] (s_18_19.center) -- (r_18_19.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_12_15) at (141mm,225mm) {};
\draw[->] (s_12_15.center) -- (r_12_15.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_12_14) at (151mm,210mm) {};
\draw[->] (s_12_14.center) -- (r_12_14.center); 
\node[draw,fill=blue!20,minimum size=10mm] (r_12_13) at (161mm,195mm) {};
\draw[->] (s_12_13.center) -- (r_12_13.center);