### 2013-06-07 graphs/visu_gruyere_binarytree.tex

\node (core_0_label) at (0mm,0mm) {core 00};
\node (core_1_label) at (0mm,5mm) {core 01};
\node (core_2_label) at (0mm,10mm) {core 02};
\node (core_3_label) at (0mm,15mm) {core 03};
\node (core_4_label) at (0mm,20mm) {core 04};
\node (core_5_label) at (0mm,25mm) {core 05};
\node (core_6_label) at (0mm,30mm) {core 06};
\node (core_7_label) at (0mm,35mm) {core 07};
\node (core_8_label) at (0mm,40mm) {core 08};
\node (core_9_label) at (0mm,45mm) {core 09};
\node (core_10_label) at (0mm,50mm) {core 10};
\node (core_11_label) at (0mm,55mm) {core 11};
\node (core_12_label) at (0mm,60mm) {core 12};
\node (core_13_label) at (0mm,65mm) {core 13};
\node (core_14_label) at (0mm,70mm) {core 14};
\node (core_15_label) at (0mm,75mm) {core 15};
\node (core_16_label) at (0mm,80mm) {core 16};
\node (core_17_label) at (0mm,85mm) {core 17};
\node (core_18_label) at (0mm,90mm) {core 18};
\node (core_19_label) at (0mm,95mm) {core 19};
\node (core_20_label) at (0mm,100mm) {core 20};
\node (core_21_label) at (0mm,105mm) {core 21};
\node (core_22_label) at (0mm,110mm) {core 22};
\node (core_23_label) at (0mm,115mm) {core 23};
\node (core_24_label) at (0mm,120mm) {core 24};
\node (core_25_label) at (0mm,125mm) {core 25};
\node (core_26_label) at (0mm,130mm) {core 26};
\node (core_27_label) at (0mm,135mm) {core 27};
\node (core_28_label) at (0mm,140mm) {core 28};
\node (core_29_label) at (0mm,145mm) {core 29};
\node (core_30_label) at (0mm,150mm) {core 30};
\node (core_31_label) at (0mm,155mm) {core 31};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_0_1) at (20mm,0mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_0_2) at (24mm,0mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_0_1) at (24mm,5mm) {};
\draw[->] ($(s_0_1.east)-(1mm,0mm)$) -- ($(r_0_1.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_1_3) at (28mm,5mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_1_4) at (32mm,5mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_0_2) at (28mm,10mm) {};
\draw[->] ($(s_0_2.east)-(1mm,0mm)$) -- ($(r_0_2.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_2_6) at (32mm,10mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_2_5) at (36mm,10mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_1_3) at (32mm,15mm) {};
\draw[->] ($(s_1_3.east)-(1mm,0mm)$) -- ($(r_1_3.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_3_8) at (36mm,15mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_3_7) at (40mm,15mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_1_4) at (40mm,20mm) {};
\draw[->,semithick,color=red] ($(s_1_4.east)-(1mm,0mm)$) -- ($(r_1_4.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_4_9) at (44mm,20mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_4_10) at (48mm,20mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_2_6) at (40mm,30mm) {};
\draw[->,semithick,color=red] ($(s_2_6.east)-(1mm,0mm)$) -- ($(r_2_6.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_6_14) at (44mm,30mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_6_13) at (48mm,30mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_2_5) at (44mm,25mm) {};
\draw[->,semithick,color=red] ($(s_2_5.east)-(1mm,0mm)$) -- ($(r_2_5.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_5_12) at (48mm,25mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_5_11) at (52mm,25mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_3_7) at (48mm,35mm) {};
\draw[->,semithick,color=red] ($(s_3_7.east)-(1mm,0mm)$) -- ($(r_3_7.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_7_15) at (52mm,35mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_7_16) at (56mm,35mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_3_8) at (48mm,40mm) {};
\draw[->,semithick,color=red] ($(s_3_8.east)-(1mm,0mm)$) -- ($(r_3_8.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_8_18) at (52mm,40mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_8_17) at (56mm,40mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_4_9) at (52mm,45mm) {};
\draw[->,semithick,color=red] ($(s_4_9.east)-(1mm,0mm)$) -- ($(r_4_9.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_9_19) at (56mm,45mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_9_20) at (60mm,45mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_4_10) at (56mm,50mm) {};
\draw[->,semithick,color=red] ($(s_4_10.east)-(1mm,0mm)$) -- ($(r_4_10.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_10_22) at (60mm,50mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_10_21) at (64mm,50mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_6_14) at (56mm,70mm) {};
\draw[->,semithick,color=red] ($(s_6_14.east)-(1mm,0mm)$) -- ($(r_6_14.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_14_30) at (60mm,70mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_14_29) at (64mm,70mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_5_11) at (60mm,55mm) {};
\draw[->,semithick,color=red] ($(s_5_11.east)-(1mm,0mm)$) -- ($(r_5_11.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_11_23) at (64mm,55mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_11_24) at (68mm,55mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_5_12) at (60mm,60mm) {};
\draw[->,semithick,color=red] ($(s_5_12.east)-(1mm,0mm)$) -- ($(r_5_12.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_12_26) at (64mm,60mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_12_25) at (68mm,60mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_6_13) at (60mm,65mm) {};
\draw[->,semithick,color=red] ($(s_6_13.east)-(1mm,0mm)$) -- ($(r_6_13.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_13_28) at (64mm,65mm) {};
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_13_27) at (68mm,65mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_7_15) at (64mm,75mm) {};
\draw[->,semithick,color=red] ($(s_7_15.east)-(1mm,0mm)$) -- ($(r_7_15.west)+(1mm,0mm)$); 
\node[draw,fill=red!20,minimum width=4mm, minimum height=3mm,anchor=west] (s_15_31) at (68mm,75mm) {};
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_14_30) at (68mm,150mm) {};
\draw[->,semithick,color=red] ($(s_14_30.east)-(1mm,0mm)$) -- ($(r_14_30.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_8_18) at (68mm,90mm) {};
\draw[->,semithick,color=red] ($(s_8_18.east)-(1mm,0mm)$) -- ($(r_8_18.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_7_16) at (68mm,80mm) {};
\draw[->,semithick,color=red] ($(s_7_16.east)-(1mm,0mm)$) -- ($(r_7_16.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_12_26) at (72mm,130mm) {};
\draw[->,semithick,color=red] ($(s_12_26.east)-(1mm,0mm)$) -- ($(r_12_26.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_13_28) at (72mm,140mm) {};
\draw[->,semithick,color=red] ($(s_13_28.east)-(1mm,0mm)$) -- ($(r_13_28.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_9_19) at (72mm,95mm) {};
\draw[->,semithick,color=red] ($(s_9_19.east)-(1mm,0mm)$) -- ($(r_9_19.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_10_22) at (72mm,110mm) {};
\draw[->,semithick,color=red] ($(s_10_22.east)-(1mm,0mm)$) -- ($(r_10_22.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_9_20) at (72mm,100mm) {};
\draw[->,semithick,color=red] ($(s_9_20.east)-(1mm,0mm)$) -- ($(r_9_20.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_14_29) at (72mm,145mm) {};
\draw[->,semithick,color=red] ($(s_14_29.east)-(1mm,0mm)$) -- ($(r_14_29.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_8_17) at (72mm,85mm) {};
\draw[->,semithick,color=red] ($(s_8_17.east)-(1mm,0mm)$) -- ($(r_8_17.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_11_24) at (76mm,120mm) {};
\draw[->,semithick,color=red] ($(s_11_24.east)-(1mm,0mm)$) -- ($(r_11_24.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_10_21) at (76mm,105mm) {};
\draw[->,semithick,color=red] ($(s_10_21.east)-(1mm,0mm)$) -- ($(r_10_21.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_13_27) at (76mm,135mm) {};
\draw[->,semithick,color=red] ($(s_13_27.east)-(1mm,0mm)$) -- ($(r_13_27.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_11_23) at (76mm,115mm) {};
\draw[->,semithick,color=red] ($(s_11_23.east)-(1mm,0mm)$) -- ($(r_11_23.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_12_25) at (76mm,125mm) {};
\draw[->,semithick,color=red] ($(s_12_25.east)-(1mm,0mm)$) -- ($(r_12_25.west)+(1mm,0mm)$); 
\node[draw,fill=blue!20,minimum width=4mm, minimum height=3mm,anchor=west] (r_15_31) at (76mm,155mm) {};
\draw[->,semithick,color=red] ($(s_15_31.east)-(1mm,0mm)$) -- ($(r_15_31.west)+(1mm,0mm)$); 
\begin{pgfonlayer}{background}
\node [fit=(core_0_label) (core_1_label) (core_2_label) (core_3_label) (core_4_label) (core_5_label) (core_6_label) (core_7_label) (core_8_label) (core_9_label) (core_10_label) (core_11_label) (core_12_label) (core_13_label) (core_14_label) (core_15_label) (core_16_label) (core_17_label) (core_18_label) (core_19_label) (core_20_label) (core_21_label) (core_22_label) (core_23_label) (core_24_label) (core_25_label) (core_26_label) (core_27_label) (core_28_label) (core_29_label) (core_30_label) (core_31_label) (s_0_1) (s_0_2) (r_0_1) (s_1_3) (s_1_4) (r_0_2) (s_2_6) (s_2_5) (r_1_3) (s_3_8) (s_3_7) (r_1_4) (s_4_9) (s_4_10) (r_2_6) (s_6_14) (s_6_13) (r_2_5) (s_5_12) (s_5_11) (r_3_7) (s_7_15) (s_7_16) (r_3_8) (s_8_18) (s_8_17) (r_4_9) (s_9_19) (s_9_20) (r_4_10) (s_10_22) (s_10_21) (r_6_14) (s_14_30) (s_14_29) (r_6_13) (s_13_28) (s_13_27) (r_5_11) (s_11_23) (s_11_24) (r_5_12) (s_12_26) (s_12_25) (r_7_15) (s_15_31) (r_14_30) (r_7_16) (r_8_18) (r_9_19) (r_12_26) (r_10_22) (r_9_20) (r_14_29) (r_13_28) (r_8_17) (r_11_24) (r_12_25) (r_11_23) (r_13_27) (r_10_21) (r_15_31) (core_0_label) (core_1_label) (core_2_label) (core_3_label) (core_4_label) (core_5_label) (core_6_label) (core_7_label) (core_8_label) (core_9_label) (core_10_label) (core_11_label) (core_12_label) (core_13_label) (core_14_label) (core_15_label) (core_16_label) (core_17_label) (core_18_label) (core_19_label) (core_20_label) (core_21_label) (core_22_label) (core_23_label) (core_24_label) (core_25_label) (core_26_label) (core_27_label) (core_28_label) (core_29_label) (core_30_label) (core_31_label) (s_0_1) (s_0_2) (r_0_1) (s_1_3) (s_1_4) (r_0_2) (s_2_6) (s_2_5) (r_1_3) (s_3_8) (s_3_7) (r_1_4) (s_4_9) (s_4_10) (r_2_6) (s_6_14) (s_6_13) (r_2_5) (s_5_12) (s_5_11) (r_3_7) (s_7_15) (s_7_16) (r_3_8) (s_8_18) (s_8_17) (r_4_9) (s_9_19) (s_9_20) (r_4_10) (s_10_22) (s_10_21) (r_6_14) (s_14_30) (s_14_29) (r_5_11) (s_11_23) (s_11_24) (r_5_12) (s_12_26) (s_12_25) (r_6_13) (s_13_28) (s_13_27) (r_7_15) (s_15_31) (r_14_30) (r_8_18) (r_7_16) (r_12_26) (r_13_28) (r_9_19) (r_10_22) (r_9_20) (r_14_29) (r_8_17) (r_11_24) (r_10_21) (r_13_27) (r_11_23) (r_12_25) (r_15_31)] (allobjects) {};
\node [draw=black!50,fill=black!10,fit=(core_0_label) (core_1_label) (core_2_label) (core_3_label) (core_4_label) (core_5_label) (core_6_label) (core_7_label) (core_8_label) (core_9_label) (core_10_label) (core_11_label) (core_12_label) (core_13_label) (core_14_label) (core_15_label) (core_16_label) (core_17_label) (core_18_label) (core_19_label) (core_20_label) (core_21_label) (core_22_label) (core_23_label) (core_24_label) (core_25_label) (core_26_label) (core_27_label) (core_28_label) (core_29_label) (core_30_label) (core_31_label) (s_0_1) (s_0_2) (r_0_1) (s_1_3) (s_1_4) (r_0_2) (s_2_6) (s_2_5) (r_1_3) (s_3_8) (s_3_7) (r_1_4) (s_4_9) (s_4_10) (r_2_6) (s_6_14) (s_6_13) (r_2_5) (s_5_12) (s_5_11) (r_3_7) (s_7_15) (s_7_16) (r_3_8) (s_8_18) (s_8_17) (r_4_9) (s_9_19) (s_9_20) (r_4_10) (s_10_22) (s_10_21) (r_6_14) (s_14_30) (s_14_29) (r_6_13) (s_13_28) (s_13_27) (r_5_11) (s_11_23) (s_11_24) (r_5_12) (s_12_26) (s_12_25) (r_7_15) (s_15_31) (r_14_30) (r_7_16) (r_8_18) (r_9_19) (r_12_26) (r_10_22) (r_9_20) (r_14_29) (r_13_28) (r_8_17) (r_11_24) (r_12_25) (r_11_23) (r_13_27) (r_10_21) (r_15_31) (core_0_label) (core_1_label) (core_2_label) (core_3_label) (core_4_label) (core_5_label) (core_6_label) (core_7_label) (core_8_label) (core_9_label) (core_10_label) (core_11_label) (core_12_label) (core_13_label) (core_14_label) (core_15_label) (core_16_label) (core_17_label) (core_18_label) (core_19_label) (core_20_label) (core_21_label) (core_22_label) (core_23_label) (core_24_label) (core_25_label) (core_26_label) (core_27_label) (core_28_label) (core_29_label) (core_30_label) (core_31_label) (s_0_1) (s_0_2) (r_0_1) (s_1_3) (s_1_4) (r_0_2) (s_2_6) (s_2_5) (r_1_3) (s_3_8) (s_3_7) (r_1_4) (s_4_9) (s_4_10) (r_2_6) (s_6_14) (s_6_13) (r_2_5) (s_5_12) (s_5_11) (r_3_7) (s_7_15) (s_7_16) (r_3_8) (s_8_18) (s_8_17) (r_4_9) (s_9_19) (s_9_20) (r_4_10) (s_10_22) (s_10_21) (r_6_14) (s_14_30) (s_14_29) (r_5_11) (s_11_23) (s_11_24) (r_5_12) (s_12_26) (s_12_25) (r_6_13) (s_13_28) (s_13_27) (r_7_15) (s_15_31) (r_14_30) (r_8_18) (r_7_16) (r_12_26) (r_13_28) (r_9_19) (r_10_22) (r_9_20) (r_14_29) (r_8_17) (r_11_24) (r_10_21) (r_13_27) (r_11_23) (r_12_25) (r_15_31),scale=1.1] (bg) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_0) at (\x1,0mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_1) at (\x1,5mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_2) at (\x1,10mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_3) at (\x1,15mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_4) at (\x1,20mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_5) at (\x1,25mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_6) at (\x1,30mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_7) at (\x1,35mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_8) at (\x1,40mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_9) at (\x1,45mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_10) at (\x1,50mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_11) at (\x1,55mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_12) at (\x1,60mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_13) at (\x1,65mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_14) at (\x1,70mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_15) at (\x1,75mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_16) at (\x1,80mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_17) at (\x1,85mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_18) at (\x1,90mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_19) at (\x1,95mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_20) at (\x1,100mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_21) at (\x1,105mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_22) at (\x1,110mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_23) at (\x1,115mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_24) at (\x1,120mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_25) at (\x1,125mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_26) at (\x1,130mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_27) at (\x1,135mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_28) at (\x1,140mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_29) at (\x1,145mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_30) at (\x1,150mm) {};
\draw let \p1 = (allobjects.east) in node[] (numa_axis_31) at (\x1,155mm) {};
\node [yscale=0.85,draw=red!50,fill=red!10,fit=(core_0_label) (core_1_label) (core_2_label) (core_3_label) (s_0_1) (s_0_2) (r_0_1) (s_1_3) (s_1_4) (r_0_2) (s_2_6) (s_2_5) (r_1_3) (s_3_8) (s_3_7) (numa_axis_0.west) (core_0_label) (core_1_label) (core_2_label) (core_3_label) (s_0_1) (s_0_2) (r_0_1) (s_1_3) (s_1_4) (r_0_2) (s_2_6) (s_2_5) (r_1_3) (s_3_8) (s_3_7) (numa_axis_0.west),rounded corners] {};
\node [yscale=0.85,draw=green!50,fill=green!10,fit=(core_4_label) (core_5_label) (core_6_label) (core_7_label) (r_1_4) (s_4_9) (s_4_10) (r_2_6) (s_6_14) (s_6_13) (r_2_5) (s_5_12) (s_5_11) (r_3_7) (s_7_15) (s_7_16) (numa_axis_4.west) (core_4_label) (core_5_label) (core_6_label) (core_7_label) (r_1_4) (s_4_9) (s_4_10) (r_2_6) (s_6_14) (s_6_13) (r_2_5) (s_5_12) (s_5_11) (r_3_7) (s_7_15) (s_7_16) (numa_axis_4.west),rounded corners] {};
\node [yscale=0.85,draw=blue!50,fill=blue!10,fit=(core_8_label) (core_9_label) (core_10_label) (core_11_label) (r_3_8) (s_8_18) (s_8_17) (r_4_9) (s_9_19) (s_9_20) (r_4_10) (s_10_22) (s_10_21) (r_5_11) (s_11_23) (s_11_24) (numa_axis_8.west) (core_8_label) (core_9_label) (core_10_label) (core_11_label) (r_3_8) (s_8_18) (s_8_17) (r_4_9) (s_9_19) (s_9_20) (r_4_10) (s_10_22) (s_10_21) (r_5_11) (s_11_23) (s_11_24) (numa_axis_8.west),rounded corners] {};
\node [yscale=0.85,draw=orange!50,fill=orange!10,fit=(core_12_label) (core_13_label) (core_14_label) (core_15_label) (r_6_14) (s_14_30) (s_14_29) (r_6_13) (s_13_28) (s_13_27) (r_5_12) (s_12_26) (s_12_25) (r_7_15) (s_15_31) (numa_axis_12.west) (core_12_label) (core_13_label) (core_14_label) (core_15_label) (r_6_14) (s_14_30) (s_14_29) (r_5_12) (s_12_26) (s_12_25) (r_6_13) (s_13_28) (s_13_27) (r_7_15) (s_15_31) (numa_axis_12.west),rounded corners] {};
\node [yscale=0.85,draw=red!50,fill=red!10,fit=(core_16_label) (core_17_label) (core_18_label) (core_19_label) (r_7_16) (r_8_18) (r_9_19) (r_8_17) (numa_axis_16.west) (core_16_label) (core_17_label) (core_18_label) (core_19_label) (r_8_18) (r_7_16) (r_9_19) (r_8_17) (numa_axis_16.west),rounded corners] {};
\node [yscale=0.85,draw=green!50,fill=green!10,fit=(core_20_label) (core_21_label) (core_22_label) (core_23_label) (r_10_22) (r_9_20) (r_11_23) (r_10_21) (numa_axis_20.west) (core_20_label) (core_21_label) (core_22_label) (core_23_label) (r_10_22) (r_9_20) (r_10_21) (r_11_23) (numa_axis_20.west),rounded corners] {};
\node [yscale=0.85,draw=blue!50,fill=blue!10,fit=(core_24_label) (core_25_label) (core_26_label) (core_27_label) (r_12_26) (r_11_24) (r_12_25) (r_13_27) (numa_axis_24.west) (core_24_label) (core_25_label) (core_26_label) (core_27_label) (r_12_26) (r_11_24) (r_13_27) (r_12_25) (numa_axis_24.west),rounded corners] {};
\node [yscale=0.85,draw=orange!50,fill=orange!10,fit=(core_28_label) (core_29_label) (core_30_label) (core_31_label) (r_14_30) (r_14_29) (r_13_28) (r_15_31) (numa_axis_28.west) (core_28_label) (core_29_label) (core_30_label) (core_31_label) (r_14_30) (r_13_28) (r_14_29) (r_15_31) (numa_axis_28.west),rounded corners] {};
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,0mm) -- (\x2,0mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,5mm) -- (\x2,5mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,10mm) -- (\x2,10mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,15mm) -- (\x2,15mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,20mm) -- (\x2,20mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,25mm) -- (\x2,25mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,30mm) -- (\x2,30mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,35mm) -- (\x2,35mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,40mm) -- (\x2,40mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,45mm) -- (\x2,45mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,50mm) -- (\x2,50mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,55mm) -- (\x2,55mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,60mm) -- (\x2,60mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,65mm) -- (\x2,65mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,70mm) -- (\x2,70mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,75mm) -- (\x2,75mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,80mm) -- (\x2,80mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,85mm) -- (\x2,85mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,90mm) -- (\x2,90mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,95mm) -- (\x2,95mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,100mm) -- (\x2,100mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,105mm) -- (\x2,105mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,110mm) -- (\x2,110mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,115mm) -- (\x2,115mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,120mm) -- (\x2,120mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,125mm) -- (\x2,125mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,130mm) -- (\x2,130mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,135mm) -- (\x2,135mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,140mm) -- (\x2,140mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,145mm) -- (\x2,145mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,150mm) -- (\x2,150mm);
\draw[color=black!30] let \p1 = (core_10_label.east), \p2 = (allobjects.east) in (\x1,155mm) -- (\x2,155mm);
\node[draw=black,anchor=north,fill=black!20] at (bg.north) {Machine: gruyere, topology: binarytree};
\end{pgfonlayer}