### 2013-07-01 figs/Overview_Model_Evaluation_Parser.tex
Started to TeX the architecture figure

%\documentclass[portrait]{a0poster}
\documentclass[a4wide]{article}
\usepackage{url,color,xspace,verbatim,subfig,ctable,multirow,listings}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{rotating}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{graphics}
\usepackage{enumitem}
\usepackage{times}
\usepackage{amssymb}
\usepackage[colorlinks=true]{hyperref}
\usepackage[ruled,vlined]{algorithm2e}

% ==================================================

\graphicspath{{figs/}}
\urlstyle{sf}

% tikz stuff
\usepackage{tikz}
\usepackage{pgfplots}
% configuration
\usetikzlibrary{shapes,positioning,calc,snakes,arrows,shapes,fit,backgrounds}

\begin{document}

\begin{figure}[htb]
  \centering
  \begin{tikzpicture}[
    topo/.style={
      draw,
      minimum height=1cm,
      minimum width=3cm,
      rotate=90,
    }
    ]
    \node[draw] (model) {Model};
    \node[draw] (sim) [left=of model] {Simulator};

    \draw[fill=red!10]
        ($(sim.north)+(0,1)+0*(.5,.2)$) rectangle +(1,3)
            node[midway,rotate=90,anchor=south] {min spanning tree};
    \draw[fill=red!10]
        ($(sim.north)+(0,1)+1*(.5,.2)$) rectangle +(1,3) 
            node[midway,rotate=90,anchor=south] {binary tree};
    \draw[fill=red!10]
        ($(sim.north)+(0,1)+2*(.5,.2)$) rectangle +(1,3) 
            node[midway,rotate=90,anchor=south] {adaptive tree};
    \draw[fill=red!10]
        ($(sim.north)+(0,1)+3*(.5,.2)$) rectangle +(1,3)
            node[midway,rotate=90,anchor=south] {sequential};

    \draw[->] (model.west) -- (sim.east) node[midway,above] {input};

    \draw[->] (sim.north) -- +(0,5)
  \end{tikzpicture}
\end{figure}

\end{document}