### 2013-07-12 pitch.tex
AB pitch: Version sent to Kornilios

\documentclass{article}
\usepackage{url,color,xspace,verbatim,subfig,ctable,multirow,listings}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{txfonts}
\usepackage{rotating}
\usepackage{paralist}
\usepackage{subfig}
\usepackage{graphics}
\usepackage{enumitem}
\usepackage{times}
\usepackage{amssymb}
\usepackage[colorlinks=true]{hyperref}
\usepackage[ruled]{algorithm2e}
\usepackage[toc,page]{appendix}
\usepackage{fancyhdr}
% ==================================================

\graphicspath{{figs/}}
\urlstyle{sf}

% tikz stuff
\usepackage{tikz}
\usepackage{pgfplots}

% configuration
\usetikzlibrary{shapes,positioning,calc,snakes,arrows,shapes,fit,backgrounds}
\pgfplotsset{width=.9\linewidth}
\usepgfplotslibrary{external}
\tikzexternalize

\lstset{
  language=C,
  basicstyle=\ttfamily \small,
  flexiblecolumns=false,
  basewidth={0.5em,0.45em},
  boxpos=t,
}

\newcommand{\etal}{{\it et al.}\xspace}
\newcommand{\naive}{na\"{\i}ve\xspace}
\newcommand{\Naive}{Na\"{\i}ve\xspace}
\newcommand{\textc}[1]{{\color{gray} {\footnotesize #1}}}

\fancyhf{}
\fancyfoot[L]{\today}
\fancyfoot[R]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}
\pagestyle{fancy}

% machines
\newcommand{\ziger}{Ziger (AMD Istanbul 4x6x1)}

\definecolor{skRed}{RGB}{155,25,25}
\newcommand{\stefan}[1]{
  {\color{skRed}[{\color{red}{SK}} #1]}}

\setcounter{section}{0} % Start sections with 1, not 0
\begin{document}

\title{Adaptive broadcast tree for multicores}

% email address
\newcommand{\eaddr}{stefan.kaestle@inf.ethz.ch}
\newcommand{\email}{\href{mailto:\eaddr}{\eaddr}}

\author{Stefan Kaestle\\
  \email \\
  Systems Group, ETH Zurich}

\maketitle

% keywords: modeling, simulation, overlay networks, scheduling
\newcommand{\ska}{$\rightarrow$\xspace}

\paragraph{What do you want to enable?} 
Atomic broadcast for multicores. \stefan{Maybe even more high-level:
  Performance of distributed algorithms like replication that are
  built on top of AB}

\paragraph{What problem are you solving, and why is it hard?} 
Tuning of AB to machine characteristics. This is hard because
multicores are complex hierarchical system. Furthermore, there is a
huge variety of such machines with different characteristics
influencing performance.

\paragraph{What new idea(s) will solve the problems?}
Machine model for automatic tuning as well as use of hybrid
implementations and composition to deal with hierarchies.

\paragraph{How will you go about it?}
Online measurements to augment Model \ska evaluation of different
implementations for machine (also combination of message-passing and
shared-memory) \ska pick best fit \ska auto-tune implementation

\paragraph{How will you know and show it works?}
Compare different AB implementations. Show that performance depends on
machine. Show that our evaluation picks the best one. Verify
performance benefit for high-level application.

\paragraph{What's the related work?}
People do shared-memory, as well as message passing based
implementations. But not both simultaneously!? They also do automatic
tuning of algorithms (e.g.\ load-balancing,
parallel programs, databases according to workload, network stacks)
but not for distributed algorithms on multicores?! \stefan{Guerraoui}

\paragraph{What is your hypothesis?}
We can do better with hybrid models. Automatic tuning achieves good
performance, without programmers having to deal with complex hardware details.

\end{document}