### 2013-09-16 measurements/ziger_topologies.tex
Lot's of other things (including new results)

\begin{figure}
  \caption{topology results for \ziger}
  \label{pgfplot:20130905}
  \begin{tikzpicture}[scale=1]
    \begin{axis}[ybar,ymin=0,symbolic x coords={cluster,mst,bintree,badtree,adaptivetree},xtick=data,error bars/y dir=both,error bars/y explicit,legend style={ at={(0.5,1.03)}, anchor=south },legend columns=2,xlabel={topology},ylabel={cost [cycles]}]
    \addplot coordinates {
      (cluster,     3674.399889) +- (182.956438,182.956438)
      (mst,         3752.560889) +- (242.747418,242.747418)
      (bintree,     3776.717889) +- (277.739474,277.739474)
      (badtree,     5311.408333) +- ( 72.719583, 72.719583)
      (adaptivetree,3202.816778) +- (270.823914,270.823914)
    };
    \addplot coordinates {
      (cluster,     3546.492553)
      (mst,         3477.927649)
      (bintree,     3849.849689)
      (badtree,     6171.924742)
      (adaptivetree,3202.816778)
    };
    \legend{real hardware, simulation};
    \end{axis}
  \end{tikzpicture}
\end{figure}