### 2013-09-28 full_nos.tex
Move everything one layer up

%%
\pgfsetcolor{black}
  % Edge: 2 -- 0
  \draw [] (68.686bp,77.481bp) .. controls (64.257bp,77.481bp) and (59.669bp,77.481bp)  .. (55.241bp,77.481bp);
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (61.963bp,67.481bp) node {$10$};
  % Edge: 3 -- 0
  \draw [] (79.131bp,33.832bp) .. controls (68.76bp,42.686bp) and (55.49bp,54.014bp)  .. (45.068bp,62.911bp);
  \draw (72.099bp,60.372bp) node {$10$};
  % Edge: 1 -- 3
  \draw [] (55.234bp,19.5bp) .. controls (59.663bp,19.5bp) and (64.251bp,19.5bp)  .. (68.68bp,19.5bp);
  \draw (61.957bp,9.5bp) node {$10$};
  % Edge: 3 -- 2
  \draw [] (95.92bp,37.528bp) .. controls (95.92bp,44.442bp) and (95.92bp,52.333bp)  .. (95.92bp,59.265bp);
  \draw (105.92bp,48.396bp) node {$1$};
  % Edge: 0 -- 1
  \draw [] (28bp,59.453bp) .. controls (28bp,52.539bp) and (28bp,44.648bp)  .. (28bp,37.716bp);
  \draw (38bp,48.585bp) node {$1$};
  % Edge: 1 -- 2
  \draw [] (44.789bp,33.832bp) .. controls (55.16bp,42.686bp) and (68.43bp,54.014bp)  .. (78.852bp,62.911bp);
  \draw (71.821bp,36.372bp) node {$10$};
  % Node: 1
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (28bp,19bp) ellipse (27bp and 18bp);
  \draw (28bp,19.5bp) node {$1$};
\end{scope}
  % Node: 0
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (28bp,77bp) ellipse (27bp and 18bp);
  \draw (28bp,77.481bp) node {$0$};
\end{scope}
  % Node: 3
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (96bp,20bp) ellipse (27bp and 18bp);
  \draw (95.92bp,19.5bp) node {$3$};
\end{scope}
  % Node: 2
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (96bp,77bp) ellipse (27bp and 18bp);
  \draw (95.92bp,77.481bp) node {$2$};
\end{scope}
%