### 2013-09-28 mst_gruyere.tex

%%
\pgfsetcolor{black}
  % Edge: 0 -- 16
  \draw [] (166.21bp,141.75bp) .. controls (154.21bp,137.5bp) and (139.75bp,132.37bp)  .. (127.78bp,128.12bp);
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (142.99bp,147.94bp) node {$10$};
  % Edge: 24 -- 8
  \draw [] (397.66bp,275.03bp) .. controls (391.25bp,258.58bp) and (381.78bp,234.27bp)  .. (375.41bp,217.94bp);
  \draw (372.53bp,251.48bp) node {$10$};
  % Edge: 3 -- 0
  \draw [] (144.86bp,183.29bp) .. controls (153.27bp,177.19bp) and (163.2bp,169.98bp)  .. (171.63bp,163.85bp);
  \draw (150.24bp,162.57bp) node {$1$};
  % Edge: 13 -- 12
  \draw [] (425.39bp,40.785bp) .. controls (420.15bp,53.394bp) and (413.1bp,70.342bp)  .. (407.87bp,82.936bp);
  \draw (404.63bp,57.86bp) node {$1$};
  % Edge: 28 -- 30
  \draw [] (481.35bp,212.31bp) .. controls (490.97bp,220.54bp) and (503.03bp,230.85bp)  .. (512.69bp,239.11bp);
  \draw (488.02bp,236.71bp) node {$1$};
  % Edge: 20 -- 21
  \draw [] (201.17bp,312.44bp) .. controls (191.17bp,315.94bp) and (179.56bp,319.99bp)  .. (169.53bp,323.49bp);
  \draw (189.35bp,329.97bp) node {$1$};
  % Edge: 14 -- 12
  \draw [] (455.7bp,72.393bp) .. controls (445.22bp,77.713bp) and (432.7bp,84.068bp)  .. (422.25bp,89.375bp);
  \draw (444.97bp,92.884bp) node {$1$};
  % Edge: 24 -- 25
  \draw [] (427.6bp,302.81bp) .. controls (438.25bp,307.43bp) and (450.85bp,312.89bp)  .. (461.46bp,317.49bp);
  \draw (449.53bp,298.15bp) node {$1$};
  % Edge: 23 -- 20
  \draw [] (239.16bp,369.78bp) .. controls (236.13bp,355.61bp) and (231.94bp,336.05bp)  .. (228.93bp,321.97bp);
  \draw (245.04bp,343.88bp) node {$1$};
  % Edge: 10 -- 8
  \draw [] (392.35bp,177.85bp) .. controls (389.69bp,180.39bp) and (386.93bp,183.02bp)  .. (384.27bp,185.56bp);
  \draw (379.31bp,171.71bp) node {$1$};
  % Edge: 28 -- 29
  \draw [] (480.24bp,183.31bp) .. controls (489.9bp,174.24bp) and (502.25bp,162.63bp)  .. (511.95bp,153.51bp);
  \draw (487.09bp,158.41bp) node {$1$};
  % Edge: 9 -- 8
  \draw [] (385.56bp,225.98bp) .. controls (383.61bp,223.03bp) and (381.57bp,219.97bp)  .. (379.61bp,217.03bp);
  \draw (371.59bp,228.5bp) node {$1$};
  % Edge: 26 -- 24
  \draw [] (434.35bp,350.12bp) .. controls (428.09bp,338.07bp) and (419.78bp,322.07bp)  .. (413.53bp,310.04bp);
  \draw (435.94bp,324.08bp) node {$1$};
  % Edge: 8 -- 11
  \draw [] (363.03bp,182.61bp) .. controls (360.85bp,175.62bp) and (358.35bp,167.62bp)  .. (356.17bp,160.64bp);
  \draw (347.6bp,174.63bp) node {$1$};
  % Edge: 6 -- 4
  \draw [] (239.87bp,225.9bp) .. controls (242.6bp,224.69bp) and (245.39bp,223.47bp)  .. (248.12bp,222.27bp);
  \draw (247.99bp,236.08bp) node {$1$};
  % Edge: 15 -- 12
  \draw [] (380.96bp,36.652bp) .. controls (385.14bp,50.203bp) and (390.9bp,68.839bp)  .. (395.1bp,82.452bp);
  \draw (400.03bp,56.552bp) node {$1$};
  % Edge: 20 -- 4
  \draw [] (233.67bp,286.83bp) .. controls (241.82bp,270.43bp) and (254bp,245.9bp)  .. (262.13bp,229.53bp);
  \draw (261.9bp,264.18bp) node {$10$};
  % Edge: 18 -- 16
  \draw [] (85.864bp,50.894bp) .. controls (89.75bp,65.773bp) and (95.245bp,86.812bp)  .. (99.13bp,101.68bp);
  \draw (81.497bp,79.29bp) node {$1$};
  % Edge: 27 -- 24
  \draw [] (391.64bp,357.44bp) .. controls (394.41bp,343.6bp) and (398.24bp,324.53bp)  .. (401.01bp,310.7bp);
  \draw (385.32bp,332.07bp) node {$1$};
  % Edge: 2 -- 0
  \draw [] (174.87bp,98.4bp) .. controls (178bp,108.92bp) and (181.93bp,122.13bp)  .. (185.04bp,132.6bp);
  \draw (191.95bp,112.5bp) node {$1$};
  % Edge: 28 -- 8
  \draw [] (437.62bp,198.71bp) .. controls (424.53bp,199.05bp) and (408.85bp,199.46bp)  .. (395.75bp,199.8bp);
  \draw (416.68bp,189.25bp) node {$10$};
  % Edge: 19 -- 16
  \draw [] (55.352bp,98.839bp) .. controls (63.514bp,102.34bp) and (72.694bp,106.28bp)  .. (80.851bp,109.78bp);
  \draw (63.102bp,116.31bp) node {$1$};
  % Edge: 4 -- 0
  \draw [] (252.83bp,198.58bp) .. controls (239.51bp,188.29bp) and (221.39bp,174.3bp)  .. (208.08bp,164.03bp);
  \draw (221.46bp,193.3bp) node {$10$};
  % Edge: 8 -- 12
  \draw [] (374.34bp,182.6bp) .. controls (380.17bp,164.37bp) and (389.15bp,136.27bp)  .. (394.96bp,118.1bp);
  \draw (398.65bp,154.35bp) node {$10$};
  % Edge: 20 -- 22
  \draw [] (215.78bp,321.25bp) .. controls (209.51bp,332.8bp) and (201.28bp,347.96bp)  .. (195.03bp,359.48bp);
  \draw (193.4bp,334.37bp) node {$1$};
  % Edge: 1 -- 0
  \draw [] (205.21bp,80.92bp) .. controls (201.98bp,95.938bp) and (197.42bp,117.17bp)  .. (194.19bp,132.19bp);
  \draw (188.7bp,104.55bp) node {$1$};
  % Edge: 4 -- 8
  \draw [] (297.43bp,209.11bp) .. controls (311.25bp,207.44bp) and (328.07bp,205.41bp)  .. (341.89bp,203.74bp);
  \draw (318.66bp,195.43bp) node {$10$};
  % Edge: 17 -- 16
  \draw [] (48.717bp,151.05bp) .. controls (59.414bp,144.95bp) and (72.429bp,137.53bp)  .. (83.123bp,131.43bp);
  \draw (71.92bp,153.24bp) node {$1$};
  % Edge: 4 -- 7
  \draw [] (268.29bp,194.2bp) .. controls (267.36bp,187.11bp) and (266.29bp,178.99bp)  .. (265.36bp,171.91bp);
  \draw (256.82bp,183.06bp) node {$1$};
  % Edge: 5 -- 4
  \draw [] (275.2bp,254.93bp) .. controls (274.38bp,247.23bp) and (273.43bp,238.29bp)  .. (272.61bp,230.59bp);
  \draw (263.9bp,242.76bp) node {$1$};
  % Edge: 28 -- 31
  \draw [] (492bp,197.27bp) .. controls (500.87bp,197.04bp) and (510.71bp,196.77bp)  .. (519.58bp,196.54bp);
  \draw (505.79bp,206.91bp) node {$1$};
  % Node: 24
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (405bp,293bp) ellipse (27bp and 18bp);
  \draw (404.59bp,292.83bp) node {$24$};
\end{scope}
  % Node: 25
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (484bp,327bp) ellipse (27bp and 18bp);
  \draw (484.14bp,327.33bp) node {$25$};
\end{scope}
  % Node: 26
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (443bp,367bp) ellipse (27bp and 18bp);
  \draw (443.33bp,367.42bp) node {$26$};
\end{scope}
  % Node: 27
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (388bp,375bp) ellipse (27bp and 18bp);
  \draw (388.04bp,375.34bp) node {$27$};
\end{scope}
  % Node: 20
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (225bp,304bp) ellipse (27bp and 18bp);
  \draw (225.1bp,304.09bp) node {$20$};
\end{scope}
  % Node: 21
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (145bp,332bp) ellipse (27bp and 18bp);
  \draw (145.34bp,331.94bp) node {$21$};
\end{scope}
  % Node: 22
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (186bp,377bp) ellipse (27bp and 18bp);
  \draw (185.78bp,376.53bp) node {$22$};
\end{scope}
  % Node: 23
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (243bp,388bp) ellipse (27bp and 18bp);
  \draw (243.05bp,387.99bp) node {$23$};
\end{scope}
  % Node: 28
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (465bp,198bp) ellipse (27bp and 18bp);
  \draw (464.62bp,198bp) node {$28$};
\end{scope}
  % Node: 29
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (528bp,139bp) ellipse (27bp and 18bp);
  \draw (527.83bp,138.57bp) node {$29$};
\end{scope}
  % Node: 1
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (209bp,63bp) ellipse (27bp and 18bp);
  \draw (209.11bp,62.79bp) node {$1$};
\end{scope}
  % Node: 0
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (190bp,150bp) ellipse (27bp and 18bp);
  \draw (190.3bp,150.29bp) node {$0$};
\end{scope}
  % Node: 3
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (126bp,197bp) ellipse (27bp and 18bp);
  \draw (126.37bp,196.72bp) node {$3$};
\end{scope}
  % Node: 2
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (170bp,80bp) ellipse (27bp and 18bp);
  \draw (169.52bp,80.402bp) node {$2$};
\end{scope}
  % Node: 5
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (277bp,273bp) ellipse (27bp and 18bp);
  \draw (277.14bp,273.19bp) node {$5$};
\end{scope}
  % Node: 4
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (271bp,212bp) ellipse (27bp and 18bp);
  \draw (270.67bp,212.35bp) node {$4$};
\end{scope}
  % Node: 7
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (263bp,154bp) ellipse (27bp and 18bp);
  \draw (263.01bp,153.98bp) node {$7$};
\end{scope}
  % Node: 6
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (217bp,236bp) ellipse (27bp and 18bp);
  \draw (217.18bp,235.87bp) node {$6$};
\end{scope}
  % Node: 9
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (397bp,243bp) ellipse (27bp and 18bp);
  \draw (396.64bp,242.62bp) node {$9$};
\end{scope}
  % Node: 8
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (369bp,201bp) ellipse (27bp and 18bp);
  \draw (368.62bp,200.5bp) node {$8$};
\end{scope}
  % Node: 11
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (351bp,143bp) ellipse (27bp and 18bp);
  \draw (350.65bp,142.96bp) node {$11$};
\end{scope}
  % Node: 10
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (408bp,163bp) ellipse (27bp and 18bp);
  \draw (407.85bp,163.05bp) node {$10$};
\end{scope}
  % Node: 13
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (433bp,23bp) ellipse (27bp and 18bp);
  \draw (432.66bp,23.301bp) node {$13$};
\end{scope}
  % Node: 12
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (401bp,100bp) ellipse (27bp and 18bp);
  \draw (400.63bp,100.35bp) node {$12$};
\end{scope}
  % Node: 15
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (376bp,19bp) ellipse (27bp and 18bp);
  \draw (375.51bp,19bp) node {$15$};
\end{scope}
  % Node: 14
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (478bp,61bp) ellipse (27bp and 18bp);
  \draw (477.53bp,61.31bp) node {$14$};
\end{scope}
  % Node: 17
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (28bp,163bp) ellipse (27bp and 18bp);
  \draw (28bp,162.87bp) node {$17$};
\end{scope}
  % Node: 16
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (104bp,120bp) ellipse (27bp and 18bp);
  \draw (103.82bp,119.62bp) node {$16$};
\end{scope}
  % Node: 19
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (32bp,89bp) ellipse (27bp and 18bp);
  \draw (32.345bp,88.971bp) node {$19$};
\end{scope}
  % Node: 18
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (81bp,33bp) ellipse (27bp and 18bp);
  \draw (81.173bp,32.931bp) node {$18$};
\end{scope}
  % Node: 31
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (547bp,196bp) ellipse (27bp and 18bp);
  \draw (546.84bp,195.81bp) node {$31$};
\end{scope}
  % Node: 30
\begin{scope}
  \definecolor{strokecol}{rgb}{0.0,0.0,0.0};
  \pgfsetstrokecolor{strokecol}
  \draw (530bp,254bp) ellipse (27bp and 18bp);
  \draw (529.59bp,253.56bp) node {$30$};
\end{scope}
%